Preparación de datos

### 2024-12-16 production/prod_prep.Rmd
change to quarto + add ola 7

file.exists(here::here("input/data-original/elsoc_long.Rdata"))

load(url("https://dataverse.harvard.edu/api/access/datafile/7245118")) #ELSOC
save(elsoc_long_2016_2022.2,file = here::here("input/data-original/elsoc_long.Rdata"))

@@

elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-999] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-888] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-777] <- NA
elsoc_long_2016_2022.2[elsoc_long_2016_2022.2 ==-666] <- NA

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2022.2 <- 
elsoc_long_2016_2022.2 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         just_pension=d02_01,
         just_salud=d02_03,
         des_perc=c18_11,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

#
sjmisc::frq(elsoc_long_2016_2022.2$just_educ)
elsoc_long_2016_2022.2$just_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$just_educ, 
                      label = "Education distributive justice")

#
elsoc_long_2016_2022.2$just_pension <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$just_pension, 
                      label = "Pension distributive justice")

#
elsoc_long_2016_2022.2$just_salud <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$just_salud, 
                      label = "Health distributive justice")

#
elsoc_long_2016_2022.2$des_perc <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$des_perc, 
                      label = "Income differences are too large")

#
sjmisc::frq(elsoc_long_2016_2022.2$merit_effort)
elsoc_long_2016_2022.2$merit_effort <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$merit_effort, 
                      label = "People are rewarded for their efforts")

#
sjmisc::frq(elsoc_long_2016_2022.2$merit_talent)
elsoc_long_2016_2022.2$merit_talent <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$merit_talent, 
                      label = "People are rewarded for their intelligence")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_family)
elsoc_long_2016_2022.2$ahead_family <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_family, 
                      label = "Coming from a wealthy family")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_educ)
elsoc_long_2016_2022.2$ahead_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_educ, 
                      label = "Having a good level of education")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_ambition)
elsoc_long_2016_2022.2$ahead_ambition <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_ambition, 
                      label = "Having ambition")

#
sjmisc::frq(elsoc_long_2016_2022.2$ahead_hardwork)
elsoc_long_2016_2022.2$ahead_hardwork <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ahead_hardwork, 
                      label = "Hard work")

for (i in c("just_educ", "just_pension", "just_salud", "des_perc",
            "merit_effort","merit_talent",
            "ahead_family","ahead_educ","ahead_ambition", "ahead_hardwork")) {
elsoc_long_2016_2022.2[[i]] <-
sjlabelled::set_labels(x = elsoc_long_2016_2022.2[[i]],
                       labels = c("Strongly disagree","Disagree",
                                  "Neither disagree nor agree",
                                  "Agree", "Strongly agree"))
}

sjmisc::frq(elsoc_long_2016_2022.2$just_educ)
sjmisc::frq(elsoc_long_2016_2022.2$merit_effort)
sjmisc::frq(elsoc_long_2016_2022.2$merit_talent)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_family)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_educ)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_ambition)
sjmisc::frq(elsoc_long_2016_2022.2$ahead_hardwork)
# Socioeconomic_________________________________________________________________

# Education_______________________________________
elsoc_long_2016_2022.2$educ <- 
  car::recode(elsoc_long_2016_2022.2$m01,
              "c(1,2,3)=1;c(4,5)=2;c(6,7)=3;c(8,9,10)=4; c(-888,-999)=NA")
elsoc_long_2016_2022.2$educ <-
  factor(elsoc_long_2016_2022.2$educ,
         labels = c("Primary","High school","Technical","Universitary"))

#reverse education, reference level is the highest level
#elsoc_long_2016_2022.2$educ <- forcats::fct_rev(elsoc_long_2016_2022.2$educ)

elsoc_long_2016_2022.2$educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$educ,
                      label = "Education")
sjmisc::frq(elsoc_long_2016_2022.2$educ)

#Recoding of education to years based on casen 2017.
elsoc_long_2016_2022.2$educyear<- as.numeric(
  car::recode(elsoc_long_2016_2022.2$m01, 
              "1=0;2=4.3;3=7.5;4=9.8;5=12.02;6=13.9;
               7=14.8;8=14.9;9=16.9;10=19.07;c(-888,-999)=NA", 
              as.numeric = T))

elsoc_long_2016_2022.2$educyear <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$educyear,
                      label = "Education in years")

class(elsoc_long_2016_2022.2$educyear)
sjmisc::frq(elsoc_long_2016_2022.2$educyear)

@@

#Impute midpoint of income ranges
elsoc_long_2016_2022.2$m30_rec <-
as.numeric(car::recode(elsoc_long_2016_2022.2$m30,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
elsoc_long_2016_2022.2$m30b_rec <-
as.numeric(car::recode(elsoc_long_2016_2022.2$m30b,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(elsoc_long_2016_2022.2$m30b_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2022.2$m29_rec <-
  as.numeric(car::recode(elsoc_long_2016_2022.2$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2022.2$m29_imp <- 
  ifelse(test = !is.na(elsoc_long_2016_2022.2$m29_rec),
         yes =  elsoc_long_2016_2022.2$m29_rec,
         no =  elsoc_long_2016_2022.2$m30_rec)
summary(elsoc_long_2016_2022.2$m29_imp)

elsoc_long_2016_2022.2$m29_imp <- 
  ifelse(test = is.na(elsoc_long_2016_2022.2$m29_imp),
         yes =  elsoc_long_2016_2022.2$m30b_rec,
         no =  elsoc_long_2016_2022.2$m29_imp)
summary(elsoc_long_2016_2022.2$m29_imp)

# deflate at each year's prices
elsoc_long_2016_2022.2$deflactor <-
  with(elsoc_long_2016_2022.2, case_when(
    ola == 2016 ~ 113.88 / 123.82,
    ola == 2017 ~ 116.46 / 123.82,
    ola == 2018 ~ 119.45 / 123.82,
    ola == 2019 ~ 123.82 / 123.82
  ))

# N Household:
elsoc_long_2016_2022.2 <-
  elsoc_long_2016_2022.2 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54))
sjmisc::frq(elsoc_long_2016_2022.2$n_hogar)

#Recode DK/DA to NA
elsoc_long_2016_2022.2$n_hogar_r<-
  car::recode(elsoc_long_2016_2022.2$n_hogar,"c(-888,-999)=NA")

# Per capita household income:
elsoc_long_2016_2022.2$ing_pc <- 
  (elsoc_long_2016_2022.2$m29_imp/elsoc_long_2016_2022.2$n_hogar_r)

elsoc_long_2016_2022.2$ing_pc <-
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ing_pc,
                      label = "Household income per capita")

sjmisc::descr(elsoc_long_2016_2022.2$ing_pc)

# Compute income quintiles
elsoc_long_2016_2022.2 <- elsoc_long_2016_2022.2 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

elsoc_long_2016_2022.2$quintil <- 
  factor(elsoc_long_2016_2022.2$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

#reverse quintile, reference level is the highest quintile
#elsoc_long_2016_2022.2$quintil <- forcats::fct_rev(elsoc_long_2016_2022.2$quintil)

elsoc_long_2016_2022.2$quintil <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$quintil,
                      label = "Household income quintile per capita")

sjmisc::frq(elsoc_long_2016_2022.2$quintil)

#include new quintile category with missing cases
elsoc_long_2016_2022.2$quintil1<-
  car::recode(elsoc_long_2016_2022.2$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#elsoc_long_2016_2022.2$quintil1 <- factor(elsoc_long_2016_2022.2$quintil1, c("Q1","Q2","Q3","Q4","Q5","QNA"))

elsoc_long_2016_2022.2$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2022.2$quintil1)

#Age Ranges
elsoc_long_2016_2022.2$edad <- 
  factor(car::recode(elsoc_long_2016_2022.2$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2022.2$edad <-
  sjlabelled::set_label(elsoc_long_2016_2022.2$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

elsoc_long_2016_2022.2$ess <- 
as.numeric(elsoc_long_2016_2022.2$d01_01)
sjmisc::frq(elsoc_long_2016_2022.2$ess)
#elsoc_long_2016_2022.2$ess <- sjmisc::rec(elsoc_long_2016_2022.2$ess,rec = "rev")
#sjmisc::frq(elsoc_long_2016_2022.2$ess)

elsoc_long_2016_2022.2$ess <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$ess,
                      label = "Subjective Social Status: individual")

sjmisc::frq(elsoc_long_2016_2022.2$ess)
# sjPlot::plot_frq(elsoc_long_2016_2022.2$ess)

# Political position ____________________________
elsoc_long_2016_2022.2$pos_id <-
factor(
  car::recode(
    elsoc_long_2016_2022.2$c15,
    "c(11,12,-888,-999)='Does not identify';c(0,1,2,3,4)='Left';
     c(5)='Center';c(6,7,8,9,10)='Right'"
  ),
  levels = c('Left', 'Center', 'Right', 'Does not identify')
)

elsoc_long_2016_2022.2$pos_id <- factor(elsoc_long_2016_2022.2$pos_id,
                                      levels = levels(elsoc_long_2016_2022.2$pos_id))

elsoc_long_2016_2022.2$pos_id <- 
sjlabelled::set_label(x = elsoc_long_2016_2022.2$pos_id, 
                      label = "Political identification")

sjmisc::frq(elsoc_long_2016_2022.2$pos_id)

@@

df_study1 <- 
elsoc_long_2016_2022.2 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         ponderador_long_total, segmento, estrato,
         sexo,
         # cuestion_mig,
         # starts_with("in_"),
         just_educ,
         just_pension,
         just_salud,
         des_perc,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         # ess_inm,
         # nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         # know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         # frien_inm_t1t4,
         pos_id,
         edad)

@@

# fix data to w01 values
df_study1_wide$edad_2 <-df_study1_wide$edad_1 #age
df_study1_wide$edad_3 <-df_study1_wide$edad_1
df_study1_wide$edad_4 <-df_study1_wide$edad_1
df_study1_wide$edad_5 <-df_study1_wide$edad_1
df_study1_wide$edad_6 <-df_study1_wide$edad_1

df_study1_wide$sexo_2 <-df_study1_wide$sexo_1 #sex
df_study1_wide$sexo_3 <-df_study1_wide$sexo_1
df_study1_wide$sexo_4 <-df_study1_wide$sexo_1
df_study1_wide$sexo_5 <-df_study1_wide$sexo_1
df_study1_wide$sexo_6 <-df_study1_wide$sexo_1

df_study1_wide$educ_2 <-df_study1_wide$educ_1 #education
df_study1_wide$educ_3 <-df_study1_wide$educ_1
df_study1_wide$educ_4 <-df_study1_wide$educ_1
df_study1_wide$educ_5 <-df_study1_wide$educ_1
df_study1_wide$educ_6 <-df_study1_wide$educ_1

df_study1_wide$educyear_2 <-df_study1_wide$educyear_1 #education years
df_study1_wide$educyear_3 <-df_study1_wide$educyear_1
df_study1_wide$educyear_4 <-df_study1_wide$educyear_1
df_study1_wide$educyear_5 <-df_study1_wide$educyear_1
df_study1_wide$educyear_6 <-df_study1_wide$educyear_1

df_study1_wide$quintil1_2 <-df_study1_wide$quintil1_1 #quintiles
df_study1_wide$quintil1_3 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_4 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_5 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_6 <-df_study1_wide$quintil1_1

df_study1_wide$ess_2 <-df_study1_wide$ess_1 # subjective status
df_study1_wide$ess_3 <-df_study1_wide$ess_1
df_study1_wide$ess_4 <-df_study1_wide$ess_1
df_study1_wide$ess_5 <-df_study1_wide$ess_1
df_study1_wide$ess_6 <-df_study1_wide$ess_1

df_study1_wide$pos_id_2 <-df_study1_wide$pos_id_1 # political position
df_study1_wide$pos_id_3 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_4 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_5 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_6 <-df_study1_wide$pos_id_1

df_study1_wide$comuna_2 <-df_study1_wide$comuna_1 #comuna
df_study1_wide$comuna_3 <-df_study1_wide$comuna_1
df_study1_wide$comuna_4 <-df_study1_wide$comuna_1
df_study1_wide$comuna_5 <-df_study1_wide$comuna_1
df_study1_wide$comuna_6 <-df_study1_wide$comuna_1

df_study1_wide$comunacod_2 <-df_study1_wide$comunacod_1 #comuna
df_study1_wide$comunacod_3 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_4 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_5 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_6 <-df_study1_wide$comunacod_1

@@

# filter the dataset for the waves 1 to 4 and 6
df_study1_long <-
df_study1_long %>% 
  filter(ola %in% c(1,2,3,4,6)) %>% 
  mutate(ola=factor(ola,levels = 1:6,labels = 1:6))
dim(df_study1_long) #check, now is OK

@@

sjPlot::view_df(df_study1_long,
                show.frq = T,show.values = T,show.na = T,show.prc = T,
                show.type = T)
#______________________________________________________________________________
# obtain the idencuesta for wave 6
ids <- 
  df_study1 %>% 
  select(idencuesta,ola) %>% 
  filter(ola==6) %>% 
  sjmisc::frq(idencuesta,show.na = F) %>% as.data.frame()

>>>>

#file.exists(here::here("input/data-original/elsoc_long.Rdata"))

#load(url("https://dataverse.harvard.edu/api/access/datafile/7245118")) #ELSOC
#save(elsoc_long_2016_2023,file = here::here("input/data-original/elsoc_long.Rdata"))

@@

elsoc_long_2016_2023[elsoc_long_2016_2023 ==-999] <- NA
elsoc_long_2016_2023[elsoc_long_2016_2023 ==-888] <- NA
elsoc_long_2016_2023[elsoc_long_2016_2023 ==-777] <- NA
elsoc_long_2016_2023[elsoc_long_2016_2023 ==-666] <- NA

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2023 <- 
elsoc_long_2016_2023 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         just_pension=d02_01,
         just_salud=d02_03,
         des_perc=c18_11,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

#
sjmisc::frq(elsoc_long_2016_2023$just_educ)
elsoc_long_2016_2023$just_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$just_educ, 
                      label = "Education distributive justice")

#
elsoc_long_2016_2023$just_pension <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$just_pension, 
                      label = "Pension distributive justice")

#
elsoc_long_2016_2023$just_salud <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$just_salud, 
                      label = "Health distributive justice")

#
elsoc_long_2016_2023$des_perc <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$des_perc, 
                      label = "Income differences are too large")

#
sjmisc::frq(elsoc_long_2016_2023$merit_effort)
elsoc_long_2016_2023$merit_effort <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$merit_effort, 
                      label = "People are rewarded for their efforts")

#
sjmisc::frq(elsoc_long_2016_2023$merit_talent)
elsoc_long_2016_2023$merit_talent <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$merit_talent, 
                      label = "People are rewarded for their intelligence")

#
sjmisc::frq(elsoc_long_2016_2023$ahead_family)
elsoc_long_2016_2023$ahead_family <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$ahead_family, 
                      label = "Coming from a wealthy family")

#
sjmisc::frq(elsoc_long_2016_2023$ahead_educ)
elsoc_long_2016_2023$ahead_educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$ahead_educ, 
                      label = "Having a good level of education")

#
sjmisc::frq(elsoc_long_2016_2023$ahead_ambition)
elsoc_long_2016_2023$ahead_ambition <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$ahead_ambition, 
                      label = "Having ambition")

#
sjmisc::frq(elsoc_long_2016_2023$ahead_hardwork)
elsoc_long_2016_2023$ahead_hardwork <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$ahead_hardwork, 
                      label = "Hard work")

for (i in c("just_educ", "just_pension", "just_salud", "des_perc",
            "merit_effort","merit_talent",
            "ahead_family","ahead_educ","ahead_ambition", "ahead_hardwork")) {
elsoc_long_2016_2023[[i]] <-
sjlabelled::set_labels(x = elsoc_long_2016_2023[[i]],
                       labels = c("Strongly disagree","Disagree",
                                  "Neither disagree nor agree",
                                  "Agree", "Strongly agree"))
}

sjmisc::frq(elsoc_long_2016_2023$just_educ)
sjmisc::frq(elsoc_long_2016_2023$merit_effort)
sjmisc::frq(elsoc_long_2016_2023$merit_talent)
sjmisc::frq(elsoc_long_2016_2023$ahead_family)
sjmisc::frq(elsoc_long_2016_2023$ahead_educ)
sjmisc::frq(elsoc_long_2016_2023$ahead_ambition)
sjmisc::frq(elsoc_long_2016_2023$ahead_hardwork)
# Socioeconomic_________________________________________________________________

# Education_______________________________________
elsoc_long_2016_2023$educ <- 
  car::recode(elsoc_long_2016_2023$m01,
              "c(1,2,3)=1;c(4,5)=2;c(6,7)=3;c(8,9,10)=4; c(-888,-999)=NA")
elsoc_long_2016_2023$educ <-
  factor(elsoc_long_2016_2023$educ,
         labels = c("Primary","High school","Technical","Universitary"))

#reverse education, reference level is the highest level
#elsoc_long_2016_2023$educ <- forcats::fct_rev(elsoc_long_2016_2023$educ)

elsoc_long_2016_2023$educ <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$educ,
                      label = "Education")
sjmisc::frq(elsoc_long_2016_2023$educ)

#Recoding of education to years based on casen 2017.
elsoc_long_2016_2023$educyear<- as.numeric(
  car::recode(elsoc_long_2016_2023$m01, 
              "1=0;2=4.3;3=7.5;4=9.8;5=12.02;6=13.9;
               7=14.8;8=14.9;9=16.9;10=19.07;c(-888,-999)=NA", 
              as.numeric = T))

elsoc_long_2016_2023$educyear <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$educyear,
                      label = "Education in years")

class(elsoc_long_2016_2023$educyear)
sjmisc::frq(elsoc_long_2016_2023$educyear)

@@

#Impute midpoint of income ranges
elsoc_long_2016_2023$m30_rec <-
as.numeric(car::recode(elsoc_long_2016_2023$m30,
           "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=2700000;NA=NA;c(-888,-999)=NA"))

#Impute midpoint of income ranges (2021)
elsoc_long_2016_2023$m30b_rec <-
as.numeric(car::recode(elsoc_long_2016_2023$m30b,
           "1=125000;2=300000;3=400000;4=575000;5=70000;NA=NA;c(-888,-999)=NA"))

sjmisc::frq(elsoc_long_2016_2023$m30b_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2023$m29_rec <-
  as.numeric(car::recode(elsoc_long_2016_2023$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2023$m29_imp <- 
  ifelse(test = !is.na(elsoc_long_2016_2023$m29_rec),
         yes =  elsoc_long_2016_2023$m29_rec,
         no =  elsoc_long_2016_2023$m30_rec)
summary(elsoc_long_2016_2023$m29_imp)

elsoc_long_2016_2023$m29_imp <- 
  ifelse(test = is.na(elsoc_long_2016_2023$m29_imp),
         yes =  elsoc_long_2016_2023$m30b_rec,
         no =  elsoc_long_2016_2023$m29_imp)
summary(elsoc_long_2016_2023$m29_imp)

# deflate at each year's prices
elsoc_long_2016_2023$deflactor <-
  with(elsoc_long_2016_2023, case_when(
    ola == 2016 ~ 113.88 / 123.82,
    ola == 2017 ~ 116.46 / 123.82,
    ola == 2018 ~ 119.45 / 123.82,
    ola == 2019 ~ 123.82 / 123.82
  ))

# N Household:
elsoc_long_2016_2023 <-
  elsoc_long_2016_2023 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54,
                            ola == 6 ~ m54,
                            ola == 7 ~ m54))
sjmisc::frq(elsoc_long_2016_2023$n_hogar)

#Recode DK/DA to NA
elsoc_long_2016_2023$n_hogar_r<-
  car::recode(elsoc_long_2016_2023$n_hogar,"c(-888,-999)=NA")

# Per capita household income:
elsoc_long_2016_2023$ing_pc <- 
  (elsoc_long_2016_2023$m29_imp/elsoc_long_2016_2023$n_hogar_r)

elsoc_long_2016_2023$ing_pc <-
sjlabelled::set_label(x = elsoc_long_2016_2023$ing_pc,
                      label = "Household income per capita")

sjmisc::descr(elsoc_long_2016_2023$ing_pc)

# Compute income quintiles
elsoc_long_2016_2023 <- elsoc_long_2016_2023 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

elsoc_long_2016_2023$quintil <- 
  factor(elsoc_long_2016_2023$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

#reverse quintile, reference level is the highest quintile
#elsoc_long_2016_2023$quintil <- forcats::fct_rev(elsoc_long_2016_2023$quintil)

elsoc_long_2016_2023$quintil <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$quintil,
                      label = "Household income quintile per capita")

sjmisc::frq(elsoc_long_2016_2023$quintil)

#include new quintile category with missing cases
elsoc_long_2016_2023$quintil1<-
  car::recode(elsoc_long_2016_2023$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

#elsoc_long_2016_2023$quintil1 <- factor(elsoc_long_2016_2023$quintil1, c("Q1","Q2","Q3","Q4","Q5","QNA"))

elsoc_long_2016_2023$quintil1 <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$quintil1,
                      label = "Household income quintile per capita (NA)") 
sjmisc::frq(elsoc_long_2016_2023$quintil1)

#Age Ranges
elsoc_long_2016_2023$edad <- 
  factor(car::recode(elsoc_long_2016_2023$m0_edad, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
elsoc_long_2016_2023$edad <-
  sjlabelled::set_label(elsoc_long_2016_2023$edad, 
                        label = c("Age groups")) 
# Other controls______________________________________________________________
# Subjective social status: individual ____________________________

elsoc_long_2016_2023$ess <- 
as.numeric(elsoc_long_2016_2023$d01_01)
sjmisc::frq(elsoc_long_2016_2023$ess)
#elsoc_long_2016_2023$ess <- sjmisc::rec(elsoc_long_2016_2023$ess,rec = "rev")
#sjmisc::frq(elsoc_long_2016_2023$ess)

elsoc_long_2016_2023$ess <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$ess,
                      label = "Subjective Social Status: individual")

sjmisc::frq(elsoc_long_2016_2023$ess)
# sjPlot::plot_frq(elsoc_long_2016_2023$ess)

# Political position ____________________________
elsoc_long_2016_2023$pos_id <-
factor(
  car::recode(
    elsoc_long_2016_2023$c15,
    "c(11,12,-888,-999)='Does not identify';c(0,1,2,3,4)='Left';
     c(5)='Center';c(6,7,8,9,10)='Right'"
  ),
  levels = c('Left', 'Center', 'Right', 'Does not identify')
)

elsoc_long_2016_2023$pos_id <- factor(elsoc_long_2016_2023$pos_id,
                                      levels = levels(elsoc_long_2016_2023$pos_id))

elsoc_long_2016_2023$pos_id <- 
sjlabelled::set_label(x = elsoc_long_2016_2023$pos_id, 
                      label = "Political identification")

sjmisc::frq(elsoc_long_2016_2023$pos_id)

@@

df_study1 <- 
elsoc_long_2016_2023 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         ponderador_long_total, segmento, estrato,
         sexo,
         # cuestion_mig,
         # starts_with("in_"),
         just_educ,
         just_pension,
         just_salud,
         des_perc,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         # ess_inm,
         # nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         # know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         # frien_inm_t1t4,
         pos_id,
         edad)

@@

# fix data to w01 values
df_study1_wide$edad_2 <-df_study1_wide$edad_1 #age
df_study1_wide$edad_3 <-df_study1_wide$edad_1
df_study1_wide$edad_4 <-df_study1_wide$edad_1
df_study1_wide$edad_5 <-df_study1_wide$edad_1
df_study1_wide$edad_6 <-df_study1_wide$edad_1
df_study1_wide$edad_7 <-df_study1_wide$edad_1

df_study1_wide$sexo_2 <-df_study1_wide$sexo_1 #sex
df_study1_wide$sexo_3 <-df_study1_wide$sexo_1
df_study1_wide$sexo_4 <-df_study1_wide$sexo_1
df_study1_wide$sexo_5 <-df_study1_wide$sexo_1
df_study1_wide$sexo_6 <-df_study1_wide$sexo_1
df_study1_wide$sexo_7 <-df_study1_wide$sexo_1

df_study1_wide$educ_2 <-df_study1_wide$educ_1 #education
df_study1_wide$educ_3 <-df_study1_wide$educ_1
df_study1_wide$educ_4 <-df_study1_wide$educ_1
df_study1_wide$educ_5 <-df_study1_wide$educ_1
df_study1_wide$educ_6 <-df_study1_wide$educ_1
df_study1_wide$educ_7 <-df_study1_wide$educ_1

df_study1_wide$educyear_2 <-df_study1_wide$educyear_1 #education years
df_study1_wide$educyear_3 <-df_study1_wide$educyear_1
df_study1_wide$educyear_4 <-df_study1_wide$educyear_1
df_study1_wide$educyear_5 <-df_study1_wide$educyear_1
df_study1_wide$educyear_6 <-df_study1_wide$educyear_1
df_study1_wide$educyear_7 <-df_study1_wide$educyear_1

df_study1_wide$quintil1_2 <-df_study1_wide$quintil1_1 #quintiles
df_study1_wide$quintil1_3 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_4 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_5 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_6 <-df_study1_wide$quintil1_1
df_study1_wide$quintil1_7 <-df_study1_wide$quintil1_1

df_study1_wide$ess_2 <-df_study1_wide$ess_1 # subjective status
df_study1_wide$ess_3 <-df_study1_wide$ess_1
df_study1_wide$ess_4 <-df_study1_wide$ess_1
df_study1_wide$ess_5 <-df_study1_wide$ess_1
df_study1_wide$ess_6 <-df_study1_wide$ess_1
df_study1_wide$ess_7 <-df_study1_wide$ess_1

df_study1_wide$pos_id_2 <-df_study1_wide$pos_id_1 # political position
df_study1_wide$pos_id_3 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_4 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_5 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_6 <-df_study1_wide$pos_id_1
df_study1_wide$pos_id_7 <-df_study1_wide$pos_id_1

df_study1_wide$comuna_2 <-df_study1_wide$comuna_1 #comuna
df_study1_wide$comuna_3 <-df_study1_wide$comuna_1
df_study1_wide$comuna_4 <-df_study1_wide$comuna_1
df_study1_wide$comuna_5 <-df_study1_wide$comuna_1
df_study1_wide$comuna_6 <-df_study1_wide$comuna_1
df_study1_wide$comuna_7 <-df_study1_wide$comuna_1

df_study1_wide$comunacod_2 <-df_study1_wide$comunacod_1 #comuna
df_study1_wide$comunacod_3 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_4 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_5 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_6 <-df_study1_wide$comunacod_1
df_study1_wide$comunacod_7 <-df_study1_wide$comunacod_1

@@

# filter the dataset for the waves 1 to 4 and 6
df_study1_long <-
df_study1_long %>% 
  filter(ola %in% c(1,2,3,4,6,7)) %>% 
  mutate(ola=factor(ola,levels = 1:7,labels = 1:7))
dim(df_study1_long) #check, now is OK

@@

sjPlot::view_df(df_study1_long,
                show.frq = T,show.values = T,show.na = T,show.prc = T,
                show.type = T)
#______________________________________________________________________________
# obtain the idencuesta for wave 7
ids <- 
  df_study1 %>% 
  select(idencuesta,ola) %>% 
  filter(ola==7) %>% 
  sjmisc::frq(idencuesta,show.na = F) %>% as.data.frame()
